### 2016-09-01 doc/precip_experiment_ms.Rmd
adding the first skeleton outline of the precip prediction analysis

---
bibliography: ~/Documents/precip_experiment/doc/precip_experiment.bib
csl: components/ecology.csl
fontsize: 12pt
geometry: margin=1in
header-includes:
- \usepackage{lineno}
- \usepackage{booktabs}
- \linenumbers
- \usepackage{setspace}
- \doublespacing
- \usepackage[documeZnt]{ragged2e}
- \usepackage{amsmath}
- \usepackage{lscape}
- \usepackage{bm}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
layout: 12pt
linkcolor: black
output:
  pdf_document:
    fig_caption: yes
urlcolor: black
---

```{r caching, include=FALSE}
library('png')
library('grid')
library("methods")
library("knitr")
#library("kfigr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = FALSE)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

```{R echo = FALSE}
# ds_stats <- readRDS('../data/temp_data/dataset_short_summary.RDS')
# cover_stats <- readRDS('../data/temp_data/cover_summary.RDS')
# eq_df <- readRDS('../data/temp_data/eq.RDS' ) 
# model <- readRDS('../analysis/model_out/lmer_clim_model.rds')
# extremes <- readRDS('../data/temp_data/extremes.RDS')
# species_summary <- readRDS('../data/temp_data/species_summary.RDS')
# cover_vs_production <- readRDS('../data/temp_data/cover_vs_production.RDS')
# production_n <- as.numeric(cover_vs_production[2])
# cover_n <- as.numeric(cover_vs_production[1])
```

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\begin{centering}
\textbf{\large{Predicting the effects of precipitation change on plant communities}}

\textsc{\small{Andrew R. Kleinhesselink\footnote{Corresponding author: arklein@aggiemail.usu.edu}\textsuperscript{1}, Peter B. Adler\textsuperscript{1} and Andrew Tredennick\textsuperscript{1}}}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}} \\

\end{centering}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

Abstract
--------

Global climate change has put the issue of generating and validating predictions at the forefront of ecology. Critical issues for ecological predictions are whether models derived from historical data can predict ecological responses to climate conditions that fall outside of the observed range of climate variation, and whether biotic interactions with other species need to be well understood in order to make accurate predictions. In this experiment we tested our ability to predict how a natural grassland plant community would respond to a multi-year climate manipulation experiment. We based our predictions on statistical models that relate annual climate and local competition to the growth, survival and recruitment of four plant species. Models for each demographic rate were fit using 20+ years of observational data collected during the middle part of the 20th century. For each demographic rate we tested three models with increasing levels of ecological complexity: a model without interactions between neighboring species; one with interactions from all species being treated equivalently; and one with species specific interactions. We then used each of these models to predict observations made from 2007 to 2016 in ambient and experimentally manipulated climate conditions. Our experiment shows that ....

*Key words: ecological prediction, climate change, sagebrush steppe, rain-out shelter, drought*

Introduction
-------------

Making successful predictions about the natural world has great practical utility but also is a sign of comprehensive and successful scientific understanding. In ecology the goal of prediction seems ever more urgent.  Every ecological system on earth is experiencing and is projecting to continue experiencing unprecedented rates of climate change driven by anthropogenic greenhouse gas emmissions. As a society our discussions of how to mitigate the causes of and adapt to the effects of climate change are explicitly dependent on ecological predictions.  However, ecological predictions of the consequences of climate change are far from simple and there have been few concrete tests of how well our predictions actually work in the field.

Outline of problem:

Proposed experiment:

In this experiment we tested our ability to predict how a natural grassland plant community would respond to an experimentally imposed multi-year climate manipulation. We based our predictions on statistical models that relate annual climate and local competition to the growth, survival and recruitment of four species. Models for each demographic rate were fit using 20+ years of observational data collected during the middle part of the 20th century. For each demographic rate we tested models with three levels of ecological complexity, one without any interactions between neighboring species, one including intra-specific interactions, and one including both intra-specific and inter-specific interactions. We then generated predictions from each of these models for the response of each species to the conditions in the modern experiment.

Materials and Methods
---------------------
#### Multi-year plant community data  
Long-term data on individual abundance and size of plants in a sagebrush steppe community have been collecting since the 1930s at the US Sheep Experiment Station in Dubois Idaho.  The plots were recensused beginning in 2007 using methods identical to the original studies. These data allow for the fitting of statistical models that estimate the effects of climate and competitive effects on individual plants in the plots.  In theory these statistical models can be used to project the effects of both a change in the competitive environment at this site--or a change in the climate.

### Rain-out shelter and irrigation experiment 
In the summer of 2011, we established eight pairs of new plots within the largest exclosure at the USSES. Areas for plot pairs were selected to be similar in physical environment and vegetation composition to the original plots located at this site.  Plot locations were determined by running eight transects across the grazing exclosure and establishing new plot pairs at a random location on each transect. We rejected areas that fell on hill slopes, areas with greater than 20% bare rock, or areas with over 10% cover of the woody shrubs Purshia tridentata or Amelanchiar utahensis. Plots within each pair were established 6 m away from one another and the corners of the plots were marked with steel bars. To get baseline pre-treatment data, we then censused the plant communities within the new plots with the pantograph technique as in the original plots.

In the fall of 2011, we randomly assigned one plot in each plot pair to recieve an experimental drought treatment and the other plot within each pair to recieve an experimental irrigation treatment.  We imposed an experimental drought on the drought plots with rain-out shelters consisting of transparent plastic gutters suspended over the plot.  The plastic gutters intercept rainfall falling on the plot and direct that water to collection tank. We used battery powered pumps to pump collected water out of the tanks and onto the paired irrigation treatments.  Pumps were switched on and off automatically with float switches located in the tanks.  The float switches were activated when tanks contained approximately 10 l of water. Shelters covered a 5 m^2 area around the 1 m square drought plot and we pumped the collected water through sprinklers onto a equal sized area surrounding the focal irrigation plot.  The shelter gutters were spaced to cover exactly 50% of the focal drought plot, thereby imposing a 50% reduction in rainfall from ambient conditions.  Likewise irrigation plots received 150% the ambient rainfall of control plots.

Pumps were disconnected when temperatures were regularly below freezing, from approximately October to May of each year.

### Soil water model

### Estimating climate effects on vital rates

### Generating predictions

Results
-------

Discussion
-----------

### Lessons for making ecological predictions

Conclusion
----------

Acknowledgments
---------------
This work was funded by a National Science Foundation award DEB-1054040 to PB. AK was supported by an NSF Graduate Research Fellowship and the Utah State University Graduate Studies Office. We wish to thank Kevin, Joe, Marina, Britta, Alice Willson, Carlee Coleman, Bret Taylor and the US Sheep Experiment Station.

\pagebreak{}

Tables
------

\begin{table}[ht]
\centering
\caption{Summary of datasets used in the analysis}
\label{tab:T1}
\begin{tabular}{llllll}
  \toprule
  \input{ ../data/temp_data/summary_table.txt}
\end{tabular}
\end{table}

\pagebreak{}

\begin{table}[ht]
\centering
\caption{Coefficients from a linear mixed effects model fit to the annual sagebrush data (n = `r ds_stats$n_obs`). Production indicates the difference between data estimating sagebrush cover (\protect\%) and production (g per m\protect\textsuperscript{2}). Estimates for a and b correspond to the variables in the Gompertz population model described in \ref{eq1}. Climate variables are defined in the main text.} 
\label{tab:T2}
\begin{tabular}{rrrr}
  \toprule
  \input{ ../data/temp_data/model_summary.txt}
\end{tabular}
\end{table}

\pagebreak{}

Figures
-------

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{../figs/hypothesis.png}
  \caption{Hypothetical effects of interannual climate variation on species' populations across its range. The bottom axis corresponds to average climate at each site. Figures in the top row show change in cover on the y-axis and annual temperature anomalies at each site on the x-axis. Years with above average temperatures are expected to increase population size at the coldest site (A), while they decrease population size at the warmest site (C). The middle panel plots the sensitivities of each site against the average climate gradient.  We interpret a negative slope between sensitivities to annual climate variation and average climate as support for the hypothesis that temperature controls the species' distribution.}
  \label{fig:hypothesis}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{../figs/map_locations2.png}
  \caption{Map of sites with multi-year sagebrush cover data in the western USA. Point size corresponds to number of observations at each site. Gray areas show the distribution of sagebrush based on the USGS SAGEMAP dataset (http://sagemap.wr.usgs.gov/). Inset shows an example of multi-year sagebrush cover data from three monitoring plots at Camp Williams, Utah.}
  \label{fig:map}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{../figs/effects_plot.png}
  \caption{Sensitivity of sagebrush abundance to annual climate covariates plotted against average site climate. Sensitivity is defined as the log change in sagebrush abundance produced by a 10 cm increase in precipitation (A,C) or a 1\textdegree C increase in temperature (B,D). Effects below zero indicate where above average temperatures or precipitation would decrease population size, while effects above zero indicate where above average annual temperature or precipitation would increase population size. Gray areas show 95\protect\% confidence intervals.}
  \label{fig:effects}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{../figs/predict_both.png}
  \caption{Predicted effects of a short-term increase in precipitation (A) or temperature (B) on sagebrush abundance at each of the monitoring sites.  Panel A shows the predicted proportional change after four years with 10\protect\% above average precipitation plotted against site mean annual precipitation. Panel B shows the predicted proportional change after four years of 2\protect\textdegree C above average temperatures plotted against site mean annual temperature. Gray bars show bootstrapped 95\protect\% confidence intervals. Confidence intervals do not take into account uncertainty in random effects.}
  \label{fig:predictions}
\end{figure}

\pagebreak{}

References
----------