### 2017-08-04 precip/manuscript/overall_cover_change_modified.tex
adding a bunch of old edits to scripts and manuscript files

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Tue Jan 17 10:14:56 2017
\begin{table}[ht]
\centering
\caption{Treatment effects on log cover change for each species from start of experiment in 2011 to last year of experiment in 2016. Intercept gives control effects. ARTR = \textit{A. tripartita}, HECO = \textit{H. comata}, POSE = \textit{P. secunda}, PSSP = \textit{P. spicata}.} 
\label{table:coverChange}
\begin{tabular}{rllll}

  \hline
ARTR & Estimate & Std.Error & t-value & Pr($\textgreater \lvert t \rvert $)   \\ 
  (Intercept) & -0.28 & 0.46 & -0.61 & 0.55 \\ 
  TreatmentDrought & 0.94 & 0.77 & 1.21 & 0.24 \\ 
  TreatmentIrrigation & 0.89 & 0.77 & 1.15 & 0.27 \\
  \hline 
HECO & Estimate & Std.Error & t-value & Pr($\textgreater \lvert t \rvert $) \\ 
  (Intercept) & 0.40 & 0.26 & 1.54 & 0.15 \\ 
  TreatmentDrought & -3.00 & 0.58 & -5.18 & 0.001 \\ 
  TreatmentIrrigation & -0.12 & 0.5 & -0.25 & 0.81 \\ 
    \hline

POSE & Estimate & Std.Error & t-value & Pr($\textgreater \lvert t \rvert $) \\ 
  (Intercept) & -0.72 & 0.46 & -1.57 & 0.13 \\ 
  TreatmentDrought & 0.03 & 0.82 & 0.03 & 0.97 \\ 
  TreatmentIrrigation & 1.15 & 0.78 & 1.47 & 0.16 \\ 
    \hline

PSSP & Estimate & Std.Error & t-value & Pr($\textgreater \lvert t \rvert $) \\ 
  (Intercept) & 0.02 & 0.21 & 0.09 & 0.93 \\ 
  TreatmentDrought & -0.89 & 0.38 & -2.34 & 0.03 \\ 
  TreatmentIrrigation & 0.15 & 0.38 & 0.38 & 0.70 \\ 
   \hline

\end{tabular}
\end{table}