### 2017-08-04 precip/manuscript/overall_prediction_score.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Fri Dec 23 14:08:37 2016
\begin{table}[ht]
\centering
\caption{Comparison of model predictions from climate model and no climate model for each species and vital rate.  Two prediction scores are reported, MSE and lppd. Lower MSE indicates improved predictions whereas higher lppd indicates improved predictions.  Instances where the climate model outperformed the no climate model are marked with "***" in the last column. ARTR = \textit{A. tripartita}, HECO = \textit{H. comata}, POSE = \textit{P. secunda}, PSSP = \textit{P. spicata}.} 
\label{table:overallPreds}
\begin{tabular}{lllrrrl}
  \hline
species & vital\_rate & score & climate model & baseline model & diff & improved \\ 
  \hline
ARTR & growth & lppd & -186.08 & -189.17 & 3.09 & *** \\ 
  ARTR & growth & MSE & 0.55 & 0.55 & -0.00 & *** \\ 
  ARTR & recruitment & lppd & -79.05 & -77.81 & -1.24 &  \\ 
  ARTR & recruitment & MSE & 169.86 & 9.47 & 160.39 &  \\ 
  ARTR & survival & lppd & -37.55 & -36.42 & -1.13 &  \\ 
  ARTR & survival & MSE & 0.06 & 0.06 & 0.00 &  \\ 
  HECO & growth & lppd & -475.54 & -454.36 & -21.18 &  \\ 
  HECO & growth & MSE & 1.26 & 1.18 & 0.09 &  \\ 
  HECO & recruitment & lppd & -149.43 & -151.60 & 2.17 & *** \\ 
  HECO & recruitment & MSE & 288.16 & 235.57 & 52.59 &  \\ 
  HECO & survival & lppd & -158.99 & -147.06 & -11.93 &  \\ 
  HECO & survival & MSE & 0.12 & 0.11 & 0.01 &  \\ 
  POSE & growth & lppd & -1823.71 & -1831.56 & 7.85 & *** \\ 
  POSE & growth & MSE & 1.72 & 1.73 & -0.01 & *** \\ 
  POSE & recruitment & lppd & -260.05 & -257.46 & -2.59 &  \\ 
  POSE & recruitment & MSE & 45.40 & 37.32 & 8.08 &  \\ 
  POSE & survival & lppd & -698.06 & -718.60 & 20.54 & *** \\ 
  POSE & survival & MSE & 0.14 & 0.14 & -0.00 & *** \\ 
  PSSP & growth & lppd & -1232.93 & -1237.92 & 4.99 & *** \\ 
  PSSP & growth & MSE & 1.51 & 1.51 & 0.00 &  \\ 
  PSSP & recruitment & lppd & -271.34 & -273.94 & 2.60 & *** \\ 
  PSSP & recruitment & MSE & 79.09 & 42.68 & 36.41 &  \\ 
  PSSP & survival & lppd & -332.47 & -307.26 & -25.21 &  \\ 
  PSSP & survival & MSE & 0.11 & 0.10 & 0.01 &  \\ 
   \hline
\end{tabular}
\end{table}