### 2017-08-04 precip/manuscript/pgr_predictions.tex

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Tue Jan 17 13:20:20 2017
\begin{table}[H]
\centering
\caption{MSE of predicted log cover changes and correlations between log cover changes predicted and observed. Predictions for the cover changes in the experimental plots were generated either from the climate or no climate models. Instances where the climate model made better predictions than the no climate model are indicated with the "***". ARTR = \textit{A. tripartita}, HECO = \textit{H. comata}, POSE = \textit{P. secunda}, PSSP = \textit{P. spicata}.} 
\label{table:corPGR}
\begin{tabular}{rllrrrl}
  \hline
 & species & stat & baseline model & climate model & diff & improved \\ 
  \hline
1 & ARTR & cor & 0.48 & 0.19 & -0.29 &  \\ 
  2 & ARTR & MSE & 0.30 & 0.30 & 0.00 &  \\ 
  3 & HECO & cor & 0.29 & 0.22 & -0.07 &  \\ 
  4 & HECO & MSE & 0.49 & 0.57 & 0.07 &  \\ 
  5 & POSE & cor & 0.45 & 0.53 & 0.07 & *** \\ 
  6 & POSE & MSE & 0.42 & 0.41 & -0.01 & *** \\ 
  7 & PSSP & cor & 0.36 & 0.38 & 0.03 & *** \\ 
  8 & PSSP & MSE & 0.39 & 0.39 & -0.01 & *** \\ 
   \hline
\end{tabular}
\end{table}