### 2018-01-04 removals/Manuscript/BetterThanJensen.tex
stuff

% This document is for pdflatex
\documentclass[11pt]{article}

\usepackage{amsmath,amsfonts,amssymb,graphicx,natbib,subfigure,setspace}
\usepackage[vmargin=1in,hmargin=1in]{geometry}

\usepackage{enumitem}
\setlist{topsep=.125em,itemsep=-0.15em,leftmargin=0.75cm}

\usepackage[compact,small]{titlesec}

\renewcommand{\floatpagefraction}{0.96}
\renewcommand{\topfraction}{0.99}
\renewcommand{\textfraction}{0.1}

\setlength{\parskip}{0.1in}
\setlength{\parindent}{0.0in}

\clubpenalty = 10000
\widowpenalty = 10000

\numberwithin{equation}{section}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{equation} \begin{aligned}}
\newcommand{\ea}{\end{aligned} \end{equation}}

\sloppy

\usepackage{palatino}
\renewcommand{\baselinestretch}{1.2}

\begin{document}

Here is the replacement for the Jensen correction based on Gaussian growth residuals.

After the growth model is fitted,

\begin{enumerate}
\item  Extract for each plant in the data set (indexed by $k$) 
its \texttt{logarea.t0} $z_{0,k}$, the growth residual $z_{1,k} - \hat{z}_{1,k}$ (here $z_1$ is
the plant's \texttt{logarea.t1} and $\hat{z}_1$ is the predicted logarea), and the predicted variance
in growth $V_k$.

\item Compute the set of scaled residuals $r_k = (z_{1,k} - \hat{z}_{1,k})/\sqrt{V_k}$ and store them
in a vector \texttt{scaledResiduals}.

\end{enumerate}

Then in computing the one-step-ahead predictions, for each growth model the line 
\begin{verbatim}
plants$area.pred <- plants$surv.prob*exp(plants$logarea.pred + V.pred/2)
\end{verbatim} 
is replaced by the following 
\begin{verbatim}
resid.Pred <- sqrt(V.pred)*scaledResiduals
newJensen <- mean(exp(resid.Pred))
plants$area.pred <- plants$surv.prob*exp(plants$logarea.pred)*newJensen
\end{verbatim}

\end{document}