### 2018-06-01 precip/manuscript/pgr_predictions_by_treatment.tex

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Fri Jun 01 15:59:34 2018
\begin{table}[H]
\centering
\caption{MSE of predicted log cover changes and correlations between log cover changes predicted and observed. Predictions for the cover changes in the experimental plots were generated either from the year effects or the climate models. Instances where the climate model made better predictions than the year effects model are indicated with the "***". ARTR = \textit{A. tripartita}, HECO = \textit{H. comata}, POSE = \textit{P. secunda}, PSSP = \textit{P. spicata}.} 
\label{table:corPGR}
\begin{tabular}{rlllrrrl}
  \hline
 & species & Treatment & stat & no climate model & climate model & diff & improved \\ 
  \hline
1 & ARTR & Control & cor & 0.31 & 0.00 & -0.31 &  \\ 
  2 & ARTR & Control & MSE & 0.23 & 0.31 & 0.08 &  \\ 
  3 & ARTR & Drought & cor & 0.52 & 0.19 & -0.33 &  \\ 
  4 & ARTR & Drought & MSE & 0.33 & 0.34 & 0.01 &  \\ 
  5 & ARTR & Irrigation & cor & 0.54 & -0.06 & -0.60 &  \\ 
  6 & ARTR & Irrigation & MSE & 0.29 & 0.61 & 0.32 &  \\ 
  7 & HECO & Control & cor & 0.52 & 0.30 & -0.22 &  \\ 
  8 & HECO & Control & MSE & 0.28 & 0.31 & 0.03 &  \\ 
  9 & HECO & Drought & cor & 0.07 & 0.75 & 0.68 & *** \\ 
  10 & HECO & Drought & MSE & 2.14 & 1.39 & -0.74 & *** \\ 
  11 & HECO & Irrigation & cor & 0.53 & 0.05 & -0.48 &  \\ 
  12 & HECO & Irrigation & MSE & 0.52 & 0.69 & 0.18 &  \\ 
  13 & POSE & Control & cor & 0.50 & 0.55 & 0.05 & *** \\ 
  14 & POSE & Control & MSE & 0.38 & 0.39 & 0.01 &  \\ 
  15 & POSE & Drought & cor & 0.55 & 0.77 & 0.22 & *** \\ 
  16 & POSE & Drought & MSE & 0.44 & 0.44 & -0.00 & *** \\ 
  17 & POSE & Irrigation & cor & 0.52 & 0.35 & -0.17 &  \\ 
  18 & POSE & Irrigation & MSE & 0.48 & 0.40 & -0.08 & *** \\ 
  19 & PSSP & Control & cor & 0.33 & 0.36 & 0.03 & *** \\ 
  20 & PSSP & Control & MSE & 0.35 & 0.34 & -0.01 & *** \\ 
  21 & PSSP & Drought & cor & 0.53 & 0.42 & -0.11 &  \\ 
  22 & PSSP & Drought & MSE & 0.53 & 0.53 & -0.01 & *** \\ 
  23 & PSSP & Irrigation & cor & 0.45 & 0.27 & -0.18 &  \\ 
  24 & PSSP & Irrigation & MSE & 0.39 & 0.38 & -0.01 & *** \\ 
   \hline
\end{tabular}
\end{table}